---
title: Predicting activity performance with accelerometer data
author: "datasciencestudent100"
date: "Tuesday, August 19, 2014"
output: html_document
---

Summary

With today's wearable devices, we are able to collect a large amount of data about personal activity. In this project, We will use the data provided from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict how well they perform their activity. 

More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


Load training data

```{r,echo=TRUE}

fileUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
download.file(fileUrl, destfile = "pml-training.cs")

temp <- read.csv("pml-training.cs",header = TRUE)

```


Identify and filter out non-numeric columns 

```{r}
numIndex <- sapply(temp, is.numeric)
temp2<-temp[,numIndex]

````

Include only accelerometer variables and exclude statistic variables.

```{r,echo=TRUE}

# Identify the variables using grep

acce_str <- grep("acce", colnames(temp2), value = TRUE)
total_str <- grep("total", colnames(temp2), value = TRUE)
var_str <- grep("var", colnames(temp2), value = TRUE)

# Including variables

temp2<-subset(temp2, select= names(temp2)[which(names(temp2) %in% acce_str)])

# Excludding variables

temp2<-subset(temp2, select= names(temp2)[-which(names(temp2) %in% total_str)])
temp2<-subset(temp2, select= names(temp2)[-which(names(temp2) %in% var_str)])

```


Process the training data by adding "classe" column 

```{r,echo=TRUE}

TrainDF <-data.frame(temp$classe,temp2)
names(TrainDF)[1] <- "classe"

```

Load caret library and get the subset for the training data

```{r,echo=TRUE} 

library(ggplot2); 
library(caret);  

TrainIndex <- createDataPartition(TrainDF$classe, p=0.1, list = FALSE)
subsetTrainDF <- TrainDF[TrainIndex,]

table(subsetTrainDF$classe)

```

Generate a model for the performance type using random forests method

```{r,echo=TRUE}

trainPar <- trainControl(allowParallel = TRUE, method = "cv", number = 5);
modelFit <- train(classe ~ ., data =subsetTrainDF, method="rf",
                  trainControl = trainPar, importance=TRUE);
```

Variables importance with the plot

```{r,echo=TRUE}
varImportance <- varImp(modelFit)

varImportance[[1]][1:12,]
              
plot(varImportance)

```

Get the subset for the test from training data

```{r,echo=TRUE}

subsetTesting <- TrainDF[-TrainIndex,];
                                         
subsetTesting <- subsetTesting[sample(1:nrow(subsetTesting), 1000),];
```

Cross-Validation

```{r}
errorMeasure <- confusionMatrix(subsetTesting$classe, predict(modelFit,subsetTesting));
errorMeasure
```

Load the real test data 

```{r,echo=TRUE}
fileUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(fileUrl, destfile = "pml-testing.cs")

test <- read.csv("pml-training.cs",header = TRUE)

```


Subset the real test Data

```{r,echo=TRUE}

TestDF <- subset(test, 
                 select = c(names(test)[which(names(test) %in% names(TrainDF))])
                 )
````


Random sampling 20 cases from the real test data

```{r,echo=TRUE}

sampleTest <- TestDF[sample(1:nrow(TestDF), 20,
   replace=FALSE),] 

```

Predict the sample test data

```{r,echo=TRUE}

pred <- predict(modelFit,sampleTest); 

sampleTest$predRight <- (pred == sampleTest$classe)

table(pred,sampleTest$classe)

```


